Report on Late Winter2013 Production: Image Differencing

### 2013-03-28 W13report.tex
Spellcheck.

Reference catalogs and astrometry.net indicies were constructed from
the input catalogs.

@@

The Figure \ref{fig-fpthresh} shows the 
effect of misestimating the noise in the detections for each of three different
seeing values (0.6\", 0.88\", 1.2\" respectively).  In each case the top panel shows
the number additional false positives in an LSST size CCD given a fractional misestimation
of the noise in the image and a detection threshold.  A 1\% under-estimate of the noise
corresponds to ~4 additional false positives at a threshold of $5\sigma$ in the best seeing case.  
The bottom panel shows 
the effect of over-estimating the noise in the measurement.  The effect is to reduce
the number of false positves.  A 1\% under-estimate of the noise results in 2 fewer false positives
per chip.

@@

We further investigate the variance properties of the images that are
input to the detection stage, i.e. the difference image itself in the
case of prefiltering, and the difference image correlated with its PSF
in postfiltering.  These results are summarized in
Table~\ref{tab-variance2}.  We find that, empirically, the images have
very similar noise properties at the detection stage, with the
deconvolution visit having larger variance by $1\%$ but the other
visits having essentially equal noise properties, with a bias towards
the postfiltered data having slightly higher variance (while the RMS
is small, all values are greater than 1.0).  The medians of the
variance planes tell a different story though.  For the deconvolution
visit, the median per--pixel variance is $70\%$ higher.  While this
will suppress the detections of false positives, it will also
significantly suppress the detection efficiency for true variability.
For the other visits, the prefiltered per--pixel variance is
relatively {\it higher} by $3-4\%$.  This will decrease the
sensitivity of the prefiltered data to false positives; as shown
above, the rate is indeed smaller using the prefiltering pipeline.
This analysis demonstrates that this is primarily due to an {\it
  underestimate} of the per--pixel variance in the postfilter
processing: while the prefiltered variance is a 1\% underestimate, the
postfiltered variance is a 5\% underestimate.  As outlined by
\cite{Price-Stacking}, this may be due to loss of pixel covariance
when undergoing multiple convolutions, as covariance is not currently
tracked in the DM stack.  In prefiltering, the science image undergoes
a single convolution, and the template image a single convolution.  In
postfiltering, the template image undergoes a convolution, is
subtracted from the science image, and then undergoes a second
convolution with the Psf.  As we have demonstrated here, this second
convolution is where our variance bookeeping lags the empirical
variance.

@@

In Figures~\ref{corrpre} and \ref{corrpost} we correlate statistics
calculated on the control sample of {\tt KernelCandidates} with the
numbers of false detections that come out of detection and
measurement, for the prefiltering and postfiltering cases
respectively.  We do this for statistics that represent the mean
standard deviation of the pixels in teh difference image
(KCDiffimStDev), the reduced chi2 of the pixels in the difference
image (KCDiffimChiSq), and mean squared error of the residuals
(KCDiffimMseResids).  These values are defined in
Section~\ref{sec-stats}.  There is one point in each figure from each
sensor in each of the 72 production runs per prefilter and postfilter
configuration.

@@

We intend to research additional aspects of the problem going forward,
including:
\begin{itemize}
\item How the false detection rate varies with more realistic SEDs;
\item How the false detection rate varies with airmass;
\item How the system behaves with a realistic mix of stars and galaxies;
\item How the false detection rate looks in crowded fields;
\item How to perform measurement on pre--filtered difference images;
\item Adoption of Gaussian processes for spatial modeling of the kernel;
\item Investigating the differences in measurement on positive/negative sources;
\item Developing a measurement suite for sources in prefiltered data;
\item Determing how much ``on the fly'' configuration we may do to optimize PSF--matching.
\end{itemize}

@@

\begin{figure}
\centering
\subfigure[0.6$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma0.60.eps}}\qquad
\subfigure[0.88$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma0.88.eps}}\\
\subfigure[1.2$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma1.20.eps}}
\caption{
We plot here the effect on the number of false positives as a function of detection threshold and noise misestimate.  The three panels 
are for three values of seeing.  In each panel the top pane shows what happens when the noise is under-estimated and the
bottom shows what happens to the false positive rate when the noise is over-estimated.  In all cases the counts are for
the total number of false postives in a single 4000x4072 pixel random Gaussian field.
}
\label{fig-fpthresh}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/fp_v_dsky.eps}
\caption{
Misestimation of the sky value has an effect on both the number of false positives and relative abundance of positive false positives to negative false postives.
The top pane of this figure shows the relative number of +ve/-ve (-ve/+ve) if the sky background is under (over) subtracted at a particular fractional level.  The 
bottom panel shows how the total number of false positves changes for the same misestimation.  This is not a strong function of the seeing, so only the best
seeing example is shown.
}
\label{fig-skythresh}
\end{figure}

\begin{figure}
\centering{\includegraphics[width=0.9\textwidth]{figures/corrA.eps}} \\
\caption{Correlation of the statistics calculated on the control
  sample {\tt KernelCandidates} during the model fitting with the
  number of false positives per sensor coming from
  detection and measurement, for the {\bf postfiltering} pipeline.
  Results for visit v6866601 are shown as {\it red squares}, for
  v6866602 as {\it green circles}, and for v6866603 {\it blue
    triangles}.  Three of the statistics, descibed in
  Section~\ref{sec-stats}, are presented.  They clearly correlate with
  each other.  For the deconvolution visit v6866601, the statistics
  are very poor even for the ``best'' solutions.  
}
\label{corrpost}
\end{figure}

\begin{figure}
\centering{\includegraphics[width=0.9\textwidth]{figures/corrB.eps}} \\
\caption{Correlation of the statistics calculated on the control
  sample {\tt KernelCandidates} during the model fitting with the
  number of false positives per sensor coming from detection and
  measurement, for the {\bf prefiltering} pipeline.  Results for visit
  v6866601 are shown as {\it red squares}, for v6866602 as {\it green
    circles}, and for v6866603 {\it blue triangles}.  Three of the
  statistics, descibed in Section~\ref{sec-stats}, are presented.
  They clearly correlate with each other, and also tend to correlate
  with the number of false detections (bottom row).}
\label{corrpre}
\end{figure}

>>>>

Reference catalogs and astrometry.net indices were constructed from
the input catalogs.

@@

The Figure \ref{fig-fpthresh} shows the 
effect of misestimating the noise in the detections for each of three different
seeing values (0.6\", 0.88\", 1.2\" respectively).  In each case the top panel shows
the number additional false positives in an LSST size CCD given a fractional misestimation
of the noise in the image and a detection threshold.  A 1\% under-estimate of the noise
corresponds to ~4 additional false positives at a threshold of $5\sigma$ in the best seeing case.  
The bottom panel shows 
the effect of over-estimating the noise in the measurement.  The effect is to reduce
the number of false positives.  A 1\% under-estimate of the noise results in 2 fewer false positives
per chip.

@@

We further investigate the variance properties of the images that are
input to the detection stage, i.e. the difference image itself in the
case of prefiltering, and the difference image correlated with its PSF
in postfiltering.  These results are summarized in
Table~\ref{tab-variance2}.  We find that, empirically, the images have
very similar noise properties at the detection stage, with the
deconvolution visit having larger variance by $1\%$ but the other
visits having essentially equal noise properties, with a bias towards
the postfiltered data having slightly higher variance (while the RMS
is small, all values are greater than 1.0).  The medians of the
variance planes tell a different story though.  For the deconvolution
visit, the median per--pixel variance is $70\%$ higher.  While this
will suppress the detections of false positives, it will also
significantly suppress the detection efficiency for true variability.
For the other visits, the prefiltered per--pixel variance is
relatively {\it higher} by $3-4\%$.  This will decrease the
sensitivity of the prefiltered data to false positives; as shown
above, the rate is indeed smaller using the prefiltering pipeline.
This analysis demonstrates that this is primarily due to an {\it
  underestimate} of the per--pixel variance in the postfilter
processing: while the prefiltered variance is a 1\% underestimate, the
postfiltered variance is a 5\% underestimate.  As outlined by
\cite{Price-Stacking}, this may be due to loss of pixel covariance
when undergoing multiple convolutions, as covariance is not currently
tracked in the DM stack.  In prefiltering, the science image undergoes
a single convolution, and the template image a single convolution.  In
postfiltering, the template image undergoes a convolution, is
subtracted from the science image, and then undergoes a second
convolution with the Psf.  As we have demonstrated here, this second
convolution is where our variance bookkeeping lags the empirical
variance.

@@

In Figures~\ref{corrpre} and \ref{corrpost} we correlate statistics
calculated on the control sample of {\tt KernelCandidates} with the
numbers of false detections that come out of detection and
measurement, for the prefiltering and postfiltering cases
respectively.  We do this for statistics that represent the mean
standard deviation of the pixels in the difference image
(KCDiffimStDev), the reduced chi2 of the pixels in the difference
image (KCDiffimChiSq), and mean squared error of the residuals
(KCDiffimMseResids).  These values are defined in
Section~\ref{sec-stats}.  There is one point in each figure from each
sensor in each of the 72 production runs per prefilter and postfilter
configuration.

@@

We intend to research additional aspects of the problem going forward,
including:
\begin{itemize}
\item How the false detection rate varies with more realistic SEDs;
\item How the false detection rate varies with airmass;
\item How the system behaves with a realistic mix of stars and galaxies;
\item How the false detection rate looks in crowded fields;
\item How to perform measurement on pre--filtered difference images;
\item Adoption of Gaussian processes for spatial modeling of the kernel;
\item Investigating the differences in measurement on positive/negative sources;
\item Developing a measurement suite for sources in prefiltered data;
\item Determining how much ``on the fly'' configuration we may do to optimize PSF--matching.
\end{itemize}

@@

\begin{figure}
\centering
\subfigure[0.6$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma0.60.eps}}\qquad
\subfigure[0.88$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma0.88.eps}}\\
\subfigure[1.2$^{\prime\prime}$ FWHM]{\includegraphics[width=0.45\textwidth]{figures/dfp_v_sigma1.20.eps}}
\caption{
We plot here the effect on the number of false positives as a function of detection threshold and noise misestimate.  The three panels 
are for three values of seeing.  In each panel the top pane shows what happens when the noise is under-estimated and the
bottom shows what happens to the false positive rate when the noise is over-estimated.  In all cases the counts are for
the total number of false positives in a single 4000x4072 pixel random Gaussian field.
}
\label{fig-fpthresh}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/fp_v_dsky.eps}
\caption{
Misestimation of the sky value has an effect on both the number of false positives and relative abundance of positive false positives to negative false positives.
The top pane of this figure shows the relative number of +ve/-ve (-ve/+ve) if the sky background is under (over) subtracted at a particular fractional level.  The 
bottom panel shows how the total number of false positives changes for the same misestimation.  This is not a strong function of the seeing, so only the best
seeing example is shown.
}
\label{fig-skythresh}
\end{figure}

\begin{figure}
\centering{\includegraphics[width=0.9\textwidth]{figures/corrA.eps}} \\
\caption{Correlation of the statistics calculated on the control
  sample {\tt KernelCandidates} during the model fitting with the
  number of false positives per sensor coming from
  detection and measurement, for the {\bf postfiltering} pipeline.
  Results for visit v6866601 are shown as {\it red squares}, for
  v6866602 as {\it green circles}, and for v6866603 {\it blue
    triangles}.  Three of the statistics, described in
  Section~\ref{sec-stats}, are presented.  They clearly correlate with
  each other.  For the deconvolution visit v6866601, the statistics
  are very poor even for the ``best'' solutions.  
}
\label{corrpost}
\end{figure}

\begin{figure}
\centering{\includegraphics[width=0.9\textwidth]{figures/corrB.eps}} \\
\caption{Correlation of the statistics calculated on the control
  sample {\tt KernelCandidates} during the model fitting with the
  number of false positives per sensor coming from detection and
  measurement, for the {\bf prefiltering} pipeline.  Results for visit
  v6866601 are shown as {\it red squares}, for v6866602 as {\it green
    circles}, and for v6866603 {\it blue triangles}.  Three of the
  statistics, described in Section~\ref{sec-stats}, are presented.
  They clearly correlate with each other, and also tend to correlate
  with the number of false detections (bottom row).}
\label{corrpre}
\end{figure}
